---
title: "Sorting Out the Best Extractions"
format:
  gfm:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
```

```{r}
extr_meta <- read_csv("../output/extractions_metadata.csv")
head(extr_meta)
```

```{r}
extracted <- extr_meta %>%
  filter(!is.na(extraction_id))
head(extracted, 20)

```

Qubit RNA BR Assay Kit The Qubit RNA BR (Broad-Range) Assay Kit, when used with the Qubit Fluorometer, provides an accurate and selective method for the quantitation of RNA samples. Depending on sample volume, the assay kit is designed to be accurate for initial RNA sample concentrations of 0.5 to 1,200 ng/µL, providing a detection range of 10−1,200 ng.

\*\* IF RNA QUBIT IS ENTERED AS 1,201 IT WAS TOO HIGH!\*\*

```{r}
# Make RNA average column and move to after 3rd reading `qubit_rna_3`
extracted <- extracted %>% 
  mutate(
    qubit_rna_avg = rowMeans(select(.,qubit_rna_1, qubit_rna_2, qubit_rna_3))) %>%
    relocate(qubit_rna_avg, .after = qubit_rna_3) 

# Make DNA average column and move to after 3rd reading `qubit_dna_3`   
extracted <- extracted %>% 
  mutate(   
    qubit_dna_avg = rowMeans(select(.,qubit_dna_1, qubit_dna_2, qubit_dna_3))) %>%
    relocate(qubit_dna_avg, .after = qubit_dna_3)

# Save this as a running list of extracted samples
write_csv(extracted, "../output/extracted.csv")
```

```{r}
# subset to review Qubit Scores
qubit_scores <- extracted  %>%
  select(cryo_id, qubit_dna_avg, qubit_rna_avg)

print(qubit_scores)
```

```{r}
# Create base scatterplot
rna <- ggplot(data = extracted, aes(x = extraction_id, y = qubit_rna_avg, color = cryo_id)) +
  geom_point(size = 2, shape = 1) + 
  ylab("RNA Quantity in ug/uL") + 
  geom_hline(yintercept = 20, linetype = "solid", color = "red")

# save the plot as an PNG file
ggsave("../output/rna.png", plot = rna, width = 8, height = 4, dpi = 600)

# View plot
rna
```

```{r}
dna <- ggplot(data = extracted, aes(x = extraction_id, y = qubit_dna_avg, color = cryo_id)) +
  geom_point(size = 2, shape = 3) + 
  ylab("DNA Quantity in ug/uL") + 
  geom_hline(yintercept = 20, linetype = "solid", color = "red")

# save the plot as an PNG file
ggsave("../output/dna.png", plot = dna, width = 8, height = 4, dpi = 600)

# View plot
dna
```

```{r}
rna_dna <- rna +
  geom_point(data = extracted, aes(x = extraction_id, y = qubit_dna_avg, color = cryo_id), size = 2, shape = 3) +
  scale_shape_manual(name = "Nucleotides", values = c(0, 2), labels = c("RNA", "DNA")) +
  ylab("Nucleotide Quantity in ug/uL") + 
  xlab("Extraction Batch")

# save the plot as an PNG file
ggsave("../output/rna_dna.png", plot = rna_dna, width = 8, height = 4, dpi = 600)

# View plot
rna_dna
```
