---
title: "03-normalizing"
author: "Sarah Tanja"
format:
  gfm:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Packages
```{r}
install.packages('DT')
```


## Load Packages

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(DT)
```

## Load in data

```{r}
# Import data 
extracted_slim <- read_csv("../output/extracted_slim.csv")

extracted_slim <- extracted_slim %>% 
  mutate(qubit_rna_avg = round(qubit_rna_avg, digits = 2)) %>% 
  mutate(qubit_dna_avg = round(qubit_dna_avg, digits = 2))

datatable(extracted_slim)
```

#### Make DNA and RNA dataframes

```{r}
extracted_rna <- extracted_slim %>% 
    select(-qubit_dna_avg)

extracted_dna <- extracted_slim %>% 
    select(-qubit_rna_avg)
```

## Normalizing

-   RNA quantity has to be normalized to $25ng/\mu L$ and sent in a volume of 25uL in a 96-well PCR plate (or 1.5mL tubes)
-   DNA quantity has to be normalized to $20ng/\mu L$ and sent in a volume of 30uL in a 96-well PCR plate (or )

In order to do this we have to dilute! 
Enter the handy $C1*V1 = C2*V2$ concentration-volume relationship (dilution) equation!

Knowns:
- $C1$ is the average qubit score
- $V1 = 5\mu L$ (a small but reasonably accurate volume to pipette from our eluted sample to dilute in nuclease-free water)
- $C2 = 25ng/\mu L$ for RNA
- $C2 = 20ng/\mu L$ for DNA

Here, we need to solve for $V2$, which is the total end volume for the dilution.
The amount of DNase/Rnase free water we need will be $V2 - V1$, with $V1 = 5\mu L$.

Solving for V2 (the volume of nuclease-free water)
$$
C1*V1 = C2*V2 \\
$$

$$
\frac{C1*V1}{C2} = V2 \\
$$ 

Let's write a function that calculates the volumes needed for us for each of our eluted samples

```{r}
dilute <- function(C1, C2) {
  # Initialize variables
  n <- length(C1)
  C1 <- C1
  V1 <- numeric(n)  # Initialize V1 vector
  V2 <- numeric(n)  # Initialize V2 vector
  
  # Loop through rows
  for (i in seq_along(C1)) {
    # Check for missing or invalid data
    if (C1[i] <= 0) {
      cat("Skipping row ", i, ": Invalid data\n")
      next
    }
    
    # Initialize V1 and V2 for the current row
    V1_row <- 1
    V2_row <- (C1[i] * V1_row) / C2
    
    # Loop until conditions are satisfied
    while (!is.na(V1_row) && !is.na(V2_row) && (V1_row < 1 || V2_row < 50)) {
      V1_row <- V1_row + 1
      V2_row <- (C1[i] * V1_row) / C2
      
      # Break the loop if V1 exceeds a threshold value
      if(V1_row > 49) {
        cat("Skipping row ", i, ": No valid solution found\n")
        next
      }
    }
    
    # Store optimized V1 and V2 values in the corresponding vectors
    V1[i] <- V1_row
    V2[i] <- V2_row
  }
  
# Create a data frame with C1 column
  result <- data.frame(C1 = C1)

# Add V1 and V2 columns to the result data frame
  result$V1 <- V1
  result$V2 <- V2
  result$C2 <- C1*V1/V2

  return(result)
}
```

### RNA

Use the `dilute` function, which has an output of a new dataframe
```{r}
rna_dilution <- dilute(C1 = extracted_rna$qubit_rna_avg, C2 = 26)
datatable(rna_dilution)
```

Join the rna_dilution to the extracted_rna dataframe
Where qubit_rna_avg = C1
```{r}
# rename qubit_rna_avg to C1 for the merge
diluted_rna <- extracted_rna %>%
  mutate(C1 = qubit_rna_avg) 

diluted_rna <- merge(diluted_rna, rna_dilution, by = 'C1' ) %>% 
  relocate(C1, .after = qubit_rna_avg)

diluted_rna <- diluted_rna %>% 
  mutate(V1_minus_V2 = round(V2 - V1, digits = 0)) %>% 
  relocate(V1_minus_V2, .after = V1)

datatable(diluted_rna)
```
Confirm 
Where:
V1 = eluted rna volume $\mu L$

V2 = final volume $\mu L$
C2 = calculated final concentration $ug \mu L$
C2_qubit = qubit tested final concentration $ug \mu L$

```{r}

```


```{r}

# Add volume of eluted RNA to use from each sample for dilution as a new column
extracted_rna <- extracted_rna %>%
  mutate(V1_rna_vol_uL = V1rna) %>% 
  relocate(V1_rna_vol_uL, .after = qubit_rna_avg)

# Use mutate to create a new column and calculate the final volume of eluted RNA in uL (V2) needed to normalize RNA concentration to 25ng/uL. If starting concentration is less than 25, make output 'NA'.
extracted_rna <- extracted_rna %>%
  mutate(V2_rna_vol_uL = ifelse(
    qubit_rna_avg < 25,
    NA,
    round((qubit_rna_avg * V1_rna_vol_uL) / C2rna, digits = 0)
  )) %>%
  relocate(V2_rna_vol_uL, .after = V1_rna_vol_uL)

# Use mutate to create a new column that calculates the volume of nuclease free water needed to add to the eluted RNA (V1) to normalize RNA concentration to 25ng/uL 
extracted_rna <- extracted_rna %>%
  mutate(rna_nucfreewater_vol_uL = (V2_rna_vol_uL - V1_rna_vol_uL)) %>%
  relocate(rna_nucfreewater_vol_uL, .after = V1_rna_vol_uL)
```

# Confirm final volume 
extracted_rna <- extracted_rna %>% 
  mutate(diluted_vol = rna_elut_vol_norm + rna_nucfreewater_vol_norm)

# Confirm final concentration (C2, diluted_nguL) ... should all be 25ng/uL

# C1*V1/V2 = C2

# C1 is qubit_rna_avg
# V1 is rna_elut_vol_norm
# V2 is V2rna

extracted_rna <- extracted_rna %>% 
  mutate(calc_norm_nguL = round((qubit_rna_avg * rna_elut_vol_norm) / V2rna, digits = 1))

write_csv(extracted_rna, "../output/extracted_rna_norm.csv")
```

### DNA

```{r}
# final normalized dna concentration (C2) needed in nanograms per microliter (ng/uL) 
norm_dna_ngul <- 20

# final dna volume (V2) needed in microliters (uL)
# final vol should be 30, but adding a 10uL buffer for qubit and pipetting error
norm_dna_ul <- 40

# Use mutate to create a new column and calculate the volume of eluted RNA in uL needed to normalize RNA concentration to 25ng/uL
# add but if volume is greater than 50 or less than zero is.na
extracted_norm <- extracted_norm %>% 
  mutate(dna_elut_vol_norm = round((norm_dna_ngul*norm_dna_ul)/extracted$qubit_dna_avg), digits = 0) %>%
    relocate(dna_elut_vol_norm, .after = qubit_dna_avg)

# Use mutate to create a new column that calculates the volume of nuclease free water needed to add to the eluted RNA to normalize RNA concentration to 25ng/uL in a 25uL volume
extracted_norm <- extracted_norm %>% 
  mutate(dna_nucfreewater_vol_norm = (norm_dna_ul - dna_elut_vol_norm)) %>%
    relocate(dna_nucfreewater_vol_norm, .after = dna_elut_vol_norm )

```

Gotta put parameters on this.. - if qubit RNA avereage ng/uL is less than 25 we have to filter that out and just put 'NA' - if normalized eluted volume is less than $1 \mu L$ we have to somehow flag that for serial dilution chatgpt say: You can use the mutate() function in R from the dplyr package to conditionally apply math operators and assign 'NA' to a new column based on a condition. Here's an example: library(dplyr)

# Create a sample data frame

df \<- data.frame(existing_column = c(30, 20, 40))

# Mutate a new column based on condition

df \<- mutate(df, new_column = ifelse(existing_column \>= 25, existing_column \* 2, NA))

# Print the updated data frame

print(df)

This is a problem! We can't pipette that small... our pipette only goes down to $1 \mu L$! For some of the samples we will have to do a serial dilution...

All DNA and RNA samples have at least $48 \mu L$ of eluted volume to work with.
