---
title: "03-normalizing"
author: "Sarah Tanja"
format:
  gfm:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Include packages

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
```

## Load in data

```{r}
extracted_slim <- read_csv("../output/extracted_slim.csv")
head(extracted_slim)
```

## Normalizing

-   RNA quantity has to be normalized to 25ng/uL and sent in a volume of 25uL in a 96-well PCR plate (or 1.5mL tubes)
-   DNA quantity has to be normalized to 20ng/uL and sent in a volume of 30uL in a 96-well PCR plate (or )

In order to do this we have to dilute quantities above 20ng/uL $$
C1V1 = C2V2
$$ We know $C1$ and we know what we need to turn it into $C2$ and $V2$, so we can solve for $V1$, which is the amount of eluted sample we will need. We can then find the amount of DNase/Rnase free water we need to add to the subsample volume by $V2 - V1$.

For an RNA sample where the average qubit score is 640.6... $$
C1*V1 = C2*V2 \\
where \; C1 = 640.6 ng/ \mu L \\
640.6ng/ \mu L *V1 = 25ng/ \mu L * 25 \mu L \\
V1 =  \frac{25ng/ \mu L * 25 \mu L}{640.6ng/ \mu L} \\
V1 = \frac{625}{640.6} \mu L \\
V1 = 0.97 \mu L
$$ This means we have to take just $0.97 \mu L$ of eluted RNA and dilute it in $25 \mu L - 0.97 \mu L = 24.03 \mu L$ of nuclease-free water to get an RNA concentration of $25 ng /\mu L$ in a $25 \mu L$ volume.

Let's write a code that calculates the volume for us For RNA

```{r}
# final normalized rna concentration (C2) needed in nanograms per microliter (ng/uL) 
norm_rna_ngul <- 25

# final rna volume (V2) needed in microliters (uL)
# final vol should be 25, but adding a 10uL buffer for qubit and pipetting error
norm_rna_ul <- 35

# Use mutate to create a new column and calculate the volume of eluted RNA in uL needed to normalize RNA concentration to 25ng/uL
extracted_norm <- extracted_slim %>% 
  mutate(rna_elut_vol_norm = round((norm_rna_ngul*norm_rna_ul)/extracted$qubit_rna_avg), digits = 0) %>%
    relocate(rna_elut_vol_norm, .after = qubit_rna_avg)

# Use mutate to create a new column that calculates the volume of nuclease free water needed to add to the eluted RNA to normalize RNA concentration to 25ng/uL in a 25uL volume
extracted_norm <- extracted_norm %>% 
  mutate(rna_nucfreewater_vol_norm = (norm_rna_ul - rna_elut_vol_norm)) %>%
    relocate(rna_nucfreewater_vol_norm, .after = rna_elut_vol_norm )

```

For DNA

```{r}
# final normalized dna concentration (C2) needed in nanograms per microliter (ng/uL) 
norm_dna_ngul <- 20

# final dna volume (V2) needed in microliters (uL)
# final vol should be 30, but adding a 10uL buffer for qubit and pipetting error
norm_dna_ul <- 40

# Use mutate to create a new column and calculate the volume of eluted RNA in uL needed to normalize RNA concentration to 25ng/uL
# add but if volume is greater than 50 or less than zero is.na
extracted_norm <- extracted_norm %>% 
  mutate(dna_elut_vol_norm = round((norm_dna_ngul*norm_dna_ul)/extracted$qubit_dna_avg), digits = 0) %>%
    relocate(dna_elut_vol_norm, .after = qubit_dna_avg)

# Use mutate to create a new column that calculates the volume of nuclease free water needed to add to the eluted RNA to normalize RNA concentration to 25ng/uL in a 25uL volume
extracted_norm <- extracted_norm %>% 
  mutate(dna_nucfreewater_vol_norm = (norm_dna_ul - dna_elut_vol_norm)) %>%
    relocate(dna_nucfreewater_vol_norm, .after = dna_elut_vol_norm )

```

Gotta put parameters on this.. - if qubit RNA avereage ng/uL is less than 25 we have to filter that out and just put 'NA' - if normalized eluted volume is less than $1 \mu L$ we have to somehow flag that for serial dilution chatgpt say: You can use the mutate() function in R from the dplyr package to conditionally apply math operators and assign 'NA' to a new column based on a condition. Here's an example: library(dplyr)

# Create a sample data frame

df \<- data.frame(existing_column = c(30, 20, 40))

# Mutate a new column based on condition

df \<- mutate(df, new_column = ifelse(existing_column \>= 25, existing_column \* 2, NA))

# Print the updated data frame

print(df)

This is a problem! We can't pipette that small... our pipette only goes down to $1 \mu L$! For some of the samples we will have to do a serial dilution...

All DNA and RNA samples have at least $48 \mu L$ of eluted volume to work with.
